Relationships in the extremes and their influence on competition and coexistence

### 2021-09-28 paper/Paper.Rnw

%...and variance...
Although, the body of literature on the impact of increased environmental variability has expanded, event-based studies accounted for only one-fifth of experimental climate change research as of 2006 \citep{Jentsch_2007}, the most commonly manipulated extremes being temperature and rainfall. Extreme drought and heat waves have shown to impact competitive output, rearranging species composition and dominance of grass-forb communities \citep{White_2000, White_2001, Hoover_2014}; although, complete exclusion of dominant species is uncommon \citep{Lloret_2012}. \cite{Fay_2003} experimentally manipulated rainfall for more frequent and intense precipitation while keeping constant the total volume of rain over Kansas tallgrass prairie and found that while some species had reduced fitness, overall community structure remained the same. For invasive species, increased climate and resource variability allowed for greater success of invasibility in terms of density and competitive equilibrium \citep{Parepa_2013}. 
%...and extreme events (though this blends together with changes in variance)
Although, increased variability of environmental fluctuations is a newer consideration for studying environmental changes, environmental variability has been analytically shown to be important for classical coexistence, i.e. MCT \citep{Chesson_1981, Warner_1985}. Moreover, by using MCT, increased variability has been shown to promote coexistence in plants communities, via the aforementioned storage effect \citep{Adler_2006,  Angert_2009}. This present study has applications to extreme events and climate change as we consider relationships in their extremes, potentially caused by extreme events, and their implications for competition and coexistence. 
Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in general, they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
%DAN: So what jasmin was trying to do was summarize a lot of lit on effects of climate change on coexistence, in these three parts
%(means, varinace, and extreme events).

>>>>

%...and variance...
While the body of literature on the impact of increased environmental variability has expanded, event-based studies accounted for only one-fifth of experimental climate change research as of 2006 \citep{Jentsch_2007}, the most commonly manipulated extremes being temperature and rainfall.
Despite research on climate variability in relation to competition and coexistence being only a portion of that, we review some studies here.
Extreme drought and heat waves have shown to impact competitive output, rearranging species composition and dominance of grass-forb communities \citep{White_2000, White_2001, Hoover_2014}; although complete exclusion of dominant species is uncommon \citep{Lloret_2012}. \cite{Fay_2003} experimentally manipulated rainfall for more frequent and intense precipitation while keeping constant the total volume of rain over Kansas tallgrass prairie and found that while some species had reduced fitness, overall community structure remained the same. For invasive species, increased climate and resource variability allowed for greater success of invasibility in terms of density and competitive equilibrium \citep{Parepa_2013}. 
%...and extreme events (though this blends together with changes in variance)
Although increased variability of environmental fluctuations is a newer consideration for studying environmental changes, environmental variability has been analytically shown to be important for classical coexistence, i.e. MCT \citep{Chesson_1981, Warner_1985}. Moreover, by using MCT, increased variability has been shown to promote coexistence in plants communities, via the aforementioned storage effect \citep{Adler_2006,  Angert_2009}. This present study has applications to extreme events and climate change as we consider relationships in their extremes, potentially caused by extreme events, and their implications for competition and coexistence. 
Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in general, they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
%DAN: So what jasmin was trying to do was summarize a lot of lit on effects of climate change on coexistence, in these three parts
%(means, varinace, and extreme events).
